Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-05 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs: filled xx in events description; improved prior discussion

Among our beyond-GR parameters, $\{\ell, \delta \omega_k^{(n)}\delta \tau_k^{(n)}\}$,
as already mentioned in the previous section, we hold $\{\delta \omega_k^{(n)}\}$
and $\{\delta \tau_k^{(n)}\}$ fixed to theory-specific predictions, and allow only $\ell$
to freely vary.
%
For such cases, we assume a prior uniform between appropriate ranges on $\ell$,
making sure that our marginalized posterior distributions on $\ell$ do not rail
over the maximum prior value on $\ell$.

@@

The \pSEOB{} model, as described in Sec.~\ref{sec:review_pSEOB}, is an
\emph{IMR} model which infers the properties of the underlying GW signal,
including (independently) its ringdown properties, using the Bayesian formalism
above. Naturally, the most promising candidates for our analyses are high-mass
\emph{and} loud GW observations with a significant signal-to-noise ratio (SNR)
in the post-merger part of the signal.
%
The latest LVK GW catalog~\cite{LIGOScientific:2021djp} reported 90 observed
signals not all of which are relevant for a BH ringdown analysis. In fact, in
an accompanying paper~\cite{LIGOScientific:2021sio}, the \textsc{pSEOBNRv4HM}
analysis\footnote{See, in particular, Sec.VIII A.2
in~\cite{LIGOScientific:2021sio}}, which is most similar to the \pSEOB{} model
presented in this paper, identified two events which provided the strongest
bounds on measurements of the dominant $(220)$ QNM:
GW150914~\cite{LIGOScientific:2016aoc} and
GW200129~\cite{LIGOScientific:2021djp}.
%
These two events, with a total mass of $XX \Mo$ and $XX \Mo$ respectively, are
extremely similar in their source properties. These are also two of the loudest
BBH signals observed to date with a total network SNR of XX and XX
respectively.
%
Moreover, and what is more relevant for our analysis, is their post-inspiral
(merger-ringdown) SNRs, which at XX and XX respectively (see the columns for
$\rho_{\text{post-insp}}$ in Table III of~\cite{LIGOScientific:2019fpa} and
Table IV of~\cite{LIGOScientific:2021sio}).
%
In this paper, we are going to focus on these two GW events as our probes of
the BH ringdown in modified theories of gravity.

@@

First, as we explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be considered as an effective field theory, meaning it should be
considered valid only below an energy scale, or equivalently, a length scale.
%
As a cut-off length scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon, \mathfrak{m}) = \varepsilon \, G \mathfrak{m} / c^2 \,,
\label{eq:def_cutoff}
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the mass scales involved in the problem.
%
We also momentarily restored factors of $c$ and $G$ to emphasize that $\Lambda_{\rm EFT}$ has
dimensions of length and hence can be compared to the coupling strength $\ell$.
%
In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, $\varepsilon = 1/2$ was used,
but here we consider $\varepsilon \in [0, 1]$ for generality.
%
We will say that a bound has been placed on $\ell$, if most of the support of
the posterior distribution function $P(\ell)$ is in the interval
$[0, \, \Lambda_{\rm EFT}(\varepsilon, \mathfrak{m})]$.
%
In practice, this can be quantified through the cumulative distribution function
(CDF) associated with the marginalized posterior distribution $P(\ell)$, namely
%
\begin{equation}
P(\ell \leqslant \ell_{\rm max}) = \int_{-\infty}^{\,\ell_{\rm max}} P(\ell') \, \dd \ell'\,.
\label{eq:def_cdf}
\end{equation}
%
For instance, we will demand that for a bound with 90\% credibility to be meaningfully placed on $\ell$ that
%
\begin{equation}
% \textrm{CDF}(\ell) \leqslant \textrm{CDF}(T_{\rm EFT}(\varepsilon; \, \mathfrak{m})).
P(\ell \leqslant \Lambda_{\rm EFT}) \leqslant 0.9 \,,
\quad \textrm{(EFT bound)},
\label{eq:eft_bound}
\end{equation}
%
where we let $\ell_{\rm max} = \Lambda_{\rm EFT}$ in Eq.~\eqref{eq:def_cdf}, and
likewise for other credibility percentiles.

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/edgb_cdf_varying_threshold.pdf}
\caption{(Color online).~The cumulative distribution function evaluated at the
cut-off $\Lambda_{\rm EFT}(\varepsilon,\gm)$ EdGB gravity as a function of the parameter
$\varepsilon$ for both (dashed curves) $\gm = m_2$, the secondary's source
mass, and (solid curves) $\gm = M_{f}$, the remnant's source mass, without
(black curves) and with (purple curves) linear in spin QNM corrections. The
horizontal lines lay at 0.68 and 0.90. We see that in no situation the curves pass
through the 0.68 or 0.90 lines even at $\varepsilon = 1$. This means that no bound on $\ell_{\rm GB}$
can be placed with the events we analyzed.
}
\label{fig:edgb_cdf}
\end{figure}

@@

As we emphasized in Sec.~\ref{sec:remarks}, before drawing any conclusions on
the allowed values for $\ell_{\rm GB}$ from our parameter estimations, we
must first check whether the ``EFT''~\eqref{eq:eft_bound} and
``ParSpec''~\eqref{eq:parspec_bound} bounds are satisfied.
%
We check the validity of the EFT bound in Fig.~\ref{fig:edgb_cdf}. In the top (bottom) panel we show
the cumulative density function of the $\ell_{\rm GB}$ posteriors for GW150914 (GW200129), obtained
evaluating the integral~\eqref{eq:def_cdf} with $\Lambda_{\rm EFT}(\varepsilon, \mathfrak{m})$,
with the mass scale set by the secondary's mass (i.e., $\mathfrak{m} = m_2$, dashed lines) or
the remnant's mass ($\mathfrak{m} = M_f$, solid lines), while varying $\varepsilon$ between 0 and 1.
%
For GW150914, we see that for the $N_{\rm max} = 0$ curves, that the CDF never goes
past $0.2$, regardless of the mass scale $\mathfrak{m}$ used and even at $\varepsilon =
1$. This shows that that the ``EFT bound'' given by Eq.~\eqref{eq:eft_bound} is
never met at a significant credibility and thus that we cannot place a bound on
$\ell_{\rm GB}$.
%
The situation is similar for GW200129 with $N_{\rm max} = 0$ and does not
change for either event when we add spin corrections to the EdGB QNM.
%
Taken together we are led to \emph{conclude that we cannot constrain EdGB gravity with our
present model}.
%
We summarize our findings in Table~\ref{tab:summary_edgb}.

Let us place our results in a broader context. In particular, let us contrast
them against those of Ref.~\cite{Lyu:2022gdr} which placed the bound
%
$\ell_{\rm GB} \leqslant 1.33$~km
%
using the neutron star-black hole binaries GW200105 and GW200115~\cite{LIGOScientific:2021qlt}, and
of Refs.~\cite{Nair:2019iur,Perkins:2021mhb}, which placed the bound
%
$\ell_{\rm GB} \leqslant 1.7$~km,
%
by stacking the posterior of $\ell_{\rm GB}$ from a  selection of black hole binaries from the GWTC-1 and GWTC-2 catalogs
with at least one relatively low-mass binary component~\cite{}\footnote{These bounds
are, strictly speaking, valid only when the scalar field $\varphi$ is small, i.e., $\varphi \ll 1$, and
take into consideration only the leading-order scalar field contribution arising from the dilatonic coupling.}.
%
These works use the \emph{inspiral} portion of GW event only and a constraint
on EdGB gravity can be obtained because black holes in this theory carry a
monopole scalar charge which in turn can source dipolar scalar radiation,
that affects the GW phase at $-1$PN order, relative to the dominant GR contribution, and
is proportional to the difference between the charges of binary components.
%
Now, the magnitude of the scalar charge scales roughly with the inverse of the black hole's mass squared, that is,
smaller (larger) back holes carry large (smaller) scalar charges. Qualitatively, this is because the scalar field
is sourced by the Gauss-Bonnet curvature invariant, which is large for black holes with small masses.
%
Beyond-GR corrections to the inspiral are, by construction, not present in our
waveform model and constitutes a important ingredient to be added in forthcoming
versions of the model.
%
In addition to the larger mass (which suppresses the black hole's charge), the
QNM of EdGB gravity differ very little from GR as observed in~\cite{Blazquez-Salcedo:2016enn}.
%
Hence, here we have an example of a theory in which the inspiral portion of the signal
is more constraining then the ringdown portion of the signal.
%
As we will next, the situation is the opposite for dCS gravity.

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/dcs_cdf_varying_threshold.pdf}
\caption{(Color online).~The cumulative distribution function evaluated at the
cut-off $\Lambda_{\rm EFT}(\varepsilon, \gm)$ as a function of the parameter $\varepsilon$ for both
$\gm = m_2$ (solid and dashed lines) and $\gm = M_{f}$ (dot-dashed and dotted lines) without
(blue lines) and with (orange lines) linear-in-spin QNM corrections.
The horizontal lines lay at 0.68 and 0.90.
%
We see that the CDF curves for $\gm = M_{f}$ are above 90\% for $\varepsilon =
1/2$ for both events, with and without including the $n = 1$ dCS corrections to
the dominant QNM.
}
\label{fig:dcs_cdf}
\end{figure}

We now turn our attention to dCS gravity.
%
In Fig.~\ref{fig:dCS_exec_sum} we show the marginalized posterior distribution
function of $\ell_{\rm CS}$ for both events GW150914 (top panel) and GW200129
(middle panel), with different colors distinguishing cases with and without spin corrections to
dominant $(2,2,0)$~QNM.
%
Perhaps the most salient feature of these posteriors is that they are not
peaked at $\ell_{\rm CS}$, which may mislead us to believe that we are seeing a
deviation from GR. This is not the case and the situation is that same as the
one we discussed in the previous section for EdGB gravity.
%
We verified that
%
\begin{equation}
    \gamma_{\rm CS} = \left( \frac{c^2 \ell_{\rm CS}}{G M_{\rm s}} \right)^{4}\,,
\end{equation}
%
does indeed peak at zero indicating consistency with GR, similarly to what is shown in Fig.~\ref{} for $\gamma_{\rm GB}$.
%
We also see that in both cases, that the inclusion leading-order-in-spin
correction to the mode, displaces the posteriors towards smaller values of
$\ell_{\rm CS}$. This can be seen more evidently by looking at the location of
posterior peaks.
%
Finally, in the bottom panel, we show the combined result for both events.

In Fig.~\ref{fig:dcs_cdf} we show the cumulative distribution function,
obtained from Eq.~\eqref{eq:def_cdf}, with $\ell_{\rm max} = \Lambda_{\rm EFT} (\varepsilon, \gm)$,
for both GW150914 (top panel) and GW200129 (bottom panel), where $\Lambda_{\rm EFT}$
is calculated with either $\gm = m_2$ (dashed lines) or $M_{f}$ (solid lines),
and in the range $\varepsilon \in [0,1]$.
%
For GW150914, we see that with $\gm = m_2$ that Eq.~\eqref{eq:eft_bound} is not
satisfied unless $\varepsilon \approx 0.9$ (with only $n=0$ corrections) and
$\varepsilon \approx 0.7$ (with both $n=0$ and $1$ corrections).
%
The situation is different if we use $\gm = M_f$. In this case, we find that
with or without spin corrections that Eq.~\eqref{eq:eft_bound} can be satisfied
with $\varepsilon \leqslant 1/2$, i.e., below the criteria used
Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}.
%
This means that with our model's assumptions and using the remnant's source mass $M_f$ to set the
cut-off scale that we can claim an upper bound
%
\begin{equation}
\ell_{\rm CS} \leqslant 42.6~\textrm{km}
\quad \textrm{(90\% credibility)},
\end{equation}
%
on dCS gravity and would constitute the first bound on this theory with
gravitational-wave observations alone.
%
We also note that since the CDF grows fast in approximate domain $\varepsilon = [0.2, 0.4]$,
that a stronger, albeit at lower credibility,
%
\begin{equation}
\ell_{\rm CS} \leqslant 37.4~\textrm{km}
% ~( = \Lambda_{\rm EFT}(0.3, M_{f}) )
\quad \textrm{at 68\% credibility},
\end{equation}
%
can also be placed, when $N_{\rm max}=0$.

@@

In Fig.~\ref{fig:cEFT_cdf} we show the cumulative distribution functions for
both events, calculated as already described for the EdGB and dCS cases.
%
We see that curves are very similar to those of dCS gravity for GW200129 (cf.
bottom panel in Fig.~\ref{fig:cEFT_cdf}).
%
Moreover, we find that for both events both the EFT~\eqref{eq:eft_bound} and
ParSpec~\eqref{eq:parspec_bound} bounds are satisfied for $\mathfrak{m} = M_f$
and $N_{\rm max} = 1/2$, allowing us to obtain the combined bound
%
\begin{equation}
    \ell_{\rm cEFT} \leqslant 51.3~\textrm{km},
\end{equation}
%
at 90\% credibility. As also happened for our study for dCS, the find that for
the cubic EFT that the ParSpec bound is violated by the $n = 1$ corrections to
the QNMs, implying that we cannot use the associated posterior distributions on
$\ell_{\rm cEFT}$ to draw any meaningful constraint on this parameter.

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/ceft_posteriors_combined.pdf}
\caption{(Color online).~Posterior distribution function for the coupling constant $\ell_{\rm cEFT}$ for
GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
%
In all panels, we use purple (black) lines correspond to the inclusion (or not) of the linear
in spin QNM correction.
}
\label{fig:cEFT_exec_sum}
\end{figure}

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/ceft_cdf_varying_threshold.pdf}
\caption{
}
\label{fig:cEFT_cdf}
\end{figure}

>>>>

Among our beyond-GR parameters, $\{\ell, \delta \omega_k^{(n)}\delta \tau_k^{(n)}\}$,
as already mentioned in the previous section, we hold $\{\delta \omega_k^{(n)}\}$
and $\{\delta \tau_k^{(n)}\}$ fixed to theory-specific predictions, and allow only $\ell$
to freely vary.
%
We assume an uniform prior between appropriate ranges on $\ell$,
say $\ell \in [0, \ell_{\rm max}]$.
%
The lower limit is set by the fact the modified gravity theories we consider
all have $p$ even and hence we can assume $\ell > 0$ without loss of
generality.
%
The upper limit is chosen as to make sure that our marginalized posterior
distributions on $\ell$ do not rail against $\ell_{\rm max}$.

@@

The \pSEOB{} model, as described in Sec.~\ref{sec:review_pSEOB}, is an
\emph{IMR} model which infers the properties of the underlying GW signal,
including (independently) its ringdown properties, using the Bayesian formalism
above. Naturally, the most promising candidates for our analyses are high-mass
\emph{and} loud GW observations with a significant signal-to-noise ratio (SNR)
in the post-merger part of the signal.
%
The latest LVK GW catalog~\cite{LIGOScientific:2021djp} reported 90 observed
signals not all of which are relevant for a BH ringdown analysis. In fact, in
an accompanying paper~\cite{LIGOScientific:2021sio}, the \textsc{pSEOBNRv4HM}
analysis\footnote{See, in particular, Sec.VIII A.2
in~\cite{LIGOScientific:2021sio}}, which is most similar to the \pSEOB{} model
presented in this paper, identified two events which provided the strongest
bounds on measurements of the dominant $(220)$ QNM:
GW150914~\cite{LIGOScientific:2016aoc} and
GW200129~\cite{LIGOScientific:2021djp}.
%
These two events, with a total mass of $65 \Mo$ and $63.4 \Mo$ respectively, are
extremely similar in their source properties. These are also two of the loudest
BBH signals observed to date with a total network SNR of 24 and 26.8
respectively.
%
Moreover, and what is more relevant for our analysis, is their post-inspiral
(merger-ringdown) SNRs, which at 16.1 and 16 respectively (see the columns for
$\rho_{\text{post-insp}}$ in Table III of~\cite{LIGOScientific:2019fpa} and
Table IV of~\cite{LIGOScientific:2021sio}).
%
In this paper, we are going to focus on these two GW events as our probes of
the BH ringdown in modified theories of gravity.

@@

First, as we explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be considered as an effective field theory, meaning it should be
considered valid only below an energy scale, or equivalently, a length scale.
%
As a cut-off length scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon, \mathfrak{m}) = \varepsilon \, G \mathfrak{m} / c^2 \,,
\label{eq:def_cutoff}
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the mass scales involved in the problem.
%
We also momentarily restored factors of $c$ and $G$ to emphasize that $\Lambda_{\rm EFT}$ has
dimensions of length and hence can be compared to the coupling strength $\ell$.
%
In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, $\varepsilon = 1/2$ was used,
but here we consider $\varepsilon \in [0, 1]$ for generality.
%
We will say that a bound has been placed on $\ell$, if most of the support of
the posterior distribution function $P(\ell)$ is in the interval
$[0, \, \Lambda_{\rm EFT}(\varepsilon, \mathfrak{m})]$.
%
In practice, this can be quantified through the cumulative distribution function
(CDF) associated with the marginalized posterior distribution $P(\ell)$, namely
%
\begin{equation}
P(\ell \leqslant \ell_{\rm max}) = \int_{-\infty}^{\,\ell_{\rm max}} \dd \ell'\, P(\ell') \,.
\label{eq:def_cdf}
\end{equation}
%
For instance, we will demand that for a bound with 90\% credibility to be meaningfully placed on $\ell$ that
%
\begin{equation}
% \textrm{CDF}(\ell) \leqslant \textrm{CDF}(T_{\rm EFT}(\varepsilon; \, \mathfrak{m})).
P(\ell \leqslant \Lambda_{\rm EFT}) \leqslant 0.9 \,,
\quad \textrm{(EFT bound)},
\label{eq:eft_bound}
\end{equation}
%
where we let $\ell_{\rm max} = \Lambda_{\rm EFT}$ in Eq.~\eqref{eq:def_cdf}, and
likewise for other credibility percentiles.

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/edgb_cdf_varying_threshold.pdf}
\caption{(Color online).~The CDF evaluated at the
cut-off $\Lambda_{\rm EFT}(\varepsilon,\gm)$ EdGB gravity as a function of the parameter
$\varepsilon$ for both (dashed curves) $\gm = m_2$, the secondary's source
mass, and (solid curves) $\gm = M_{f}$, the remnant's source mass, without
(black curves) and with (purple curves) linear in spin QNM corrections. The
horizontal lines lay at 0.68 and 0.90. We see that in no situation the curves pass
through the 0.68 or 0.90 lines even at $\varepsilon = 1$.
%
This means that no bound on $\ell_{\rm GB}$ can be placed with the events we
analyzed.
}
\label{fig:edgb_cdf}
\end{figure}

@@

As we emphasized in Sec.~\ref{sec:remarks}, we must first check whether the
``EFT''~\eqref{eq:eft_bound} and ``ParSpec''~\eqref{eq:parspec_bound} bounds
are satisfied, before drawing any conclusions on the allowed values for
$\ell_{\rm GB}$ from our parameter estimations
%
We check the validity of the EFT bound in Fig.~\ref{fig:edgb_cdf}. In the top (bottom) panel we show
the CDF of the $\ell_{\rm GB}$ posteriors for GW150914 (GW200129), obtained
by evaluating the integral~\eqref{eq:def_cdf} with $\ell_{\rm max} = \Lambda_{\rm EFT}(\varepsilon, \mathfrak{m})$,
with the mass scale set by the secondary's mass (i.e., $\mathfrak{m} = m_2$, dashed lines) or
the remnant's mass ($\mathfrak{m} = M_f$, solid lines), while varying $\varepsilon$ between 0 and 1.
%
For GW150914, we see that for the $N_{\rm max} = 0$ curves, that the CDF never goes
past $0.2$, regardless of the mass scale $\mathfrak{m}$ used and even at $\varepsilon =
1$, at which the EFT description of the theory would not be valid anyway.
%
This shows that that the ``EFT bound'' given by Eq.~\eqref{eq:eft_bound} is
never met at a significant credibility and thus that we cannot place a bound on
$\ell_{\rm GB}$.
%
The situation is similar for GW200129 with $N_{\rm max} = 0$ and does not
change for either event when we add spin corrections to the EdGB QNM.
%
Taken together we are led to \emph{conclude that we cannot constrain EdGB gravity with our
present model}.
%
We summarize our findings in Table~\ref{tab:summary_edgb}.

Our result contrasts with those of Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}
which relied on the BBH inspiral to constrain $\ell_{\rm GB}$, as discussed in Sec.~\ref{sec:review_edgb}.
%
We see that EdGB provides an example of a theory in which the inspiral portion
of the signal is more constraining then the ringdown portion of the signal.
%
Two reasons together can explain our negative results. First, as observed by
Ref.~\cite{Blazquez-Salcedo:2016enn}, the QNMs of EdGB BHs only differ slightly
from their Schwarzschild counterparts. Second, the larger mass $M_f$ of the
remnant BH, suppresses scalar field's charge relatively to the initial binary
components.~\hs{This is a general `limitation' of ringdown test of higher-curvature theories in present detectors.
Point this out in the conclusions after all specific theories are discussed.}
%
As we will next, the situation is the opposite for dCS gravity.

% Let us place our results in a broader context. In particular, let us contrast
% them against those of Ref.~\cite{Lyu:2022gdr} which placed the bound
% %
% $\ell_{\rm GB} \leqslant 1.33$~km
% %
% using the neutron star-black hole binaries GW200105 and GW200115~\cite{LIGOScientific:2021qlt}, and
% of Refs.~\cite{Nair:2019iur,Perkins:2021mhb}, which placed the bound
% %
% $\ell_{\rm GB} \leqslant 1.7$~km,
% %
% by stacking the posterior of $\ell_{\rm GB}$ from a  selection of black hole binaries from the GWTC-1 and GWTC-2 catalogs
% with at least one relatively low-mass binary component~\cite{}\footnote{These bounds
% are, strictly speaking, valid only when the scalar field $\varphi$ is small, i.e., $\varphi \ll 1$, and
% take into consideration only the leading-order scalar field contribution arising from the dilatonic coupling.}.
% %
% These works use the \emph{inspiral} portion of GW event only and a constraint
% on EdGB gravity can be obtained because black holes in this theory carry a
% monopole scalar charge which in turn can source dipolar scalar radiation,
% that affects the GW phase at $-1$PN order, relative to the dominant GR contribution, and
% is proportional to the difference between the charges of binary components.
% %
% Now, the magnitude of the scalar charge scales roughly with the inverse of the black hole's mass squared, that is,
% smaller (larger) back holes carry large (smaller) scalar charges. Qualitatively, this is because the scalar field
% is sourced by the Gauss-Bonnet curvature invariant, which is large for black holes with small masses.
% %
% Beyond-GR corrections to the inspiral are, by construction, not present in our
% waveform model and constitutes a important ingredient to be added in forthcoming
% versions of the model.
% %
% In addition to the larger mass (which suppresses the black hole's charge), the
% QNM of EdGB gravity differ very little from GR as observed in~\cite{Blazquez-Salcedo:2016enn}.
% %
% Hence, here we have an example of a theory in which the inspiral portion of the signal
% is more constraining then the ringdown portion of the signal.
%
% As we will next, the situation is the opposite for dCS gravity.

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/dcs_cdf_varying_threshold.pdf}
\caption{(Color online).~Similar to Fig.~\ref{fig:edgb_cdf}, but for dCS gravity.
% The cumulative distribution function evaluated at the
% cut-off $\Lambda_{\rm EFT}(\varepsilon, \gm)$ as a function of the parameter $\varepsilon$ for both
% $\gm = m_2$ (solid and dashed lines) and $\gm = M_{f}$ (dot-dashed and dotted lines) without
% (blue lines) and with (orange lines) linear-in-spin QNM corrections.
% The horizontal lines lay at 0.68 and 0.90.
%
We see that the CDF curves for $\gm = M_{f}$ are above 90\% for $\varepsilon =
1/2$ for both events, with and without including the $n = 1$ dCS corrections to
the dominant QNM.
}
\label{fig:dcs_cdf}
\end{figure}

We now consider dCS gravity.
%
In Fig.~\ref{fig:dCS_exec_sum} we show the marginalized posterior distribution
function of $\ell_{\rm CS}$ for both events GW150914 (top panel) and GW200129
(middle panel), with different colors distinguishing cases with and without spin corrections to
dominant $(2,2,0)$~QNM.
%
Perhaps the most salient feature of these posteriors is that they are not
peaked at $\ell_{\rm CS}$, which may mislead us to believe that we are seeing a
deviation from GR. This is not the case and the situation is that same as the
one we discussed in the previous section for EdGB gravity.
%
We verified that
%
\begin{equation}
    \gamma_{\rm CS} = \left( \frac{c^2 \ell_{\rm CS}}{G M_{\rm s}} \right)^{4}\,,
\end{equation}
%
does indeed peak at zero indicating consistency with GR, similarly to what is shown in Fig.~\ref{} for $\gamma_{\rm GB}$.
%
We also see that in both cases, that the inclusion leading-order-in-spin
correction to the mode, displaces the posteriors towards smaller values of
$\ell_{\rm CS}$. This can be seen more evidently by looking at the location of
posterior peaks.
%
Finally, in the bottom panel, we show the combined result for both events.

In Fig.~\ref{fig:dcs_cdf} we show the CDF
obtained from Eq.~\eqref{eq:def_cdf}, with $\ell_{\rm max} = \Lambda_{\rm EFT} (\varepsilon, \gm)$,
for both GW150914 (top panel) and GW200129 (bottom panel), where $\Lambda_{\rm EFT}$
is calculated with either $\gm = m_2$ (dashed lines) or $M_{f}$ (solid lines),
and in the range $\varepsilon \in [0,1]$.
%
For GW150914, we see that with $\gm = m_2$ that Eq.~\eqref{eq:eft_bound} is not
satisfied unless $\varepsilon \approx 0.9$ (with only $n=0$ corrections) and
$\varepsilon \approx 0.7$ (with both $n=0$ and $1$ corrections).
%
The situation is different if we use $\gm = M_f$. In this case, we find that
with or without spin corrections that Eq.~\eqref{eq:eft_bound} can be satisfied
with $\varepsilon \leqslant 1/2$, i.e., below the criteria used
Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}.
%
This means that with our model's assumptions and using the remnant's source mass $M_f$ to set the
cut-off scale that we can claim an upper bound
%
\begin{equation}
\ell_{\rm CS} \leqslant 42.6~\textrm{km}
\quad \textrm{(90\% credibility)},
\end{equation}
%
on dCS gravity and would constitute the first bound on this theory with
gravitational-wave observations alone.
%
We also note that since the CDF grows fast in approximate domain $\varepsilon = [0.2, 0.4]$,
that a stronger, albeit at lower credibility,
%
\begin{equation}
\ell_{\rm CS} \leqslant 37.4~\textrm{km}
% ~( = \Lambda_{\rm EFT}(0.3, M_{f}) )
\quad \textrm{at 68\% credibility},
\end{equation}
%
can also be placed, when $N_{\rm max}=0$.

@@

In Fig.~\ref{fig:cEFT_cdf} we show the CDF for
both events, calculated as already described for the EdGB and dCS cases.
%
We see that curves are very similar to those of dCS gravity for GW200129 (cf.
bottom panel in Fig.~\ref{fig:cEFT_cdf}).
%
Moreover, we find that for both events both the EFT~\eqref{eq:eft_bound} and
ParSpec~\eqref{eq:parspec_bound} bounds are satisfied for $\mathfrak{m} = M_f$
and $N_{\rm max} = 1/2$, allowing us to obtain the combined bound
%
\begin{equation}
    \ell_{\rm cEFT} \leqslant 51.3~\textrm{km},
\end{equation}
%
at 90\% credibility. As also happened for our study for dCS, the find that for
the cubic EFT that the ParSpec bound is violated by the $n = 1$ corrections to
the QNMs, implying that we cannot use the associated posterior distributions on
$\ell_{\rm cEFT}$ to draw any meaningful constraint on this parameter.

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/ceft_posteriors_combined.pdf}
\caption{(Color online).~Posterior distribution function for the coupling constant $\ell_{\rm cEFT}$ for
GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
%
The colors distinguish different $N_{\rm max}$ in the spin expansion.
}
\label{fig:cEFT_exec_sum}
\end{figure}

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/ceft_cdf_varying_threshold.pdf}
\caption{(Color online).~Similar to Fig.~\ref{fig:edgb_cdf}, but for the cubic EFT of GR.
We see that the CDF curves for $\gm = M_{f}$ are above 90\% for $\varepsilon =
1/2$ for both events, with and without including the $n = 1$ corrections to
the dominant QNM.
}
\label{fig:cEFT_cdf}
\end{figure}
